Open-source landscape for 3D CSEM modelling

### 2020-10-21 manuscript/Manuscript.tex
Feedback RR

The actual discretized resistivity model of the three mesh types are shown in Figure~\ref{fig:results-marlim_allmodels}.
%
\plot*{results-marlim_allmodels}{width=\textwidth}{The resistivity model as obtained from the three different discretizations using rectilinear, octree, and tetrahedral meshes (3$\times$ vertical exaggeration).}
%
The models where obtained using volume averaging (\emg3d, \simpeg) and linear interpolation (\custem, \petgem), in both cases on $\log_{10}{\rho}$ values. The actual responses at the receivers (field interpolation) are obtained by spline interpolation (\emg3d), linear interpolation (\simpeg), and through the second-order basis functions (\custem, \petgem).

@@

%% 2. Discretization %%
One of the key insights we gained from this study is that creating appropriate meshes is a difficult and time-consuming task. While there exist now several open-source 3D CSEM codes the capability for automatic meshing lacks behind and currently remains a largely manual task requiring experience in the field. It is important to state that the chosen models favour structured meshes. The first two examples can easily and accurately be represented by regular meshes. Using, for instance, a dipping layer instead of three resistive cubes would turn it around and make the model more favourable for unstructured meshes. MR3D would be ideal an scenario for unstructured meshes, as it is based on also published irregular lithological horizons. Since the corresponding resistivity model, however, was defined on a regular grid, we were forced to approximate the comparatively fine regular discretization by an unstructured one and interpolate the resistivity data for being able to apply the FE codes to this problem. We are completely aware about the ineptness of this re-approximation procedure, unless it served for the cross-validation purposes. Both cases show that automatic and adaptive meshing capabilities are needed, which will increase the flexibility of 3D CSEM modelling. Such meshing codes should also take into account the physics of CSEM computation with its diffusive behaviour. It is also important to state that the unstructured mesh used for the MR3D model incorporates the entire domain and could be used for all source and receiver positions, whereas the rectilinear and octree meshes were designed for the shown receiver and corresponding source lines.

@@

We had help from several colleagues for different tasks, and would like to thank Bane Sullivan (plotting of tetrahedra meshes in \texttt{matplotlib} through \texttt{PyVista}), Seogi Kang (input and octree mesh design for the \simpeg layer and block models), and Joseph Capriotti (octree mesh implementation and volume averaging in \simpeg). We would also like to thank Paulo Menezes for the help and explanations with regards to the Marlim R3D model and corresponding CSEM data, and for making their actual computation model available under an open-source license.

>>>>

The actual discretized resistivity model of the three mesh types are shown in Figure~\ref{fig:results-marlim_allmodels}.
%
\plot*{results-marlim_allmodels}{width=\textwidth}{The resistivity model as obtained from the three different discretizations using rectilinear, octree, and tetrahedral meshes (3$\times$ vertical exaggeration).}
%
The models where obtained using volume averaging (\emg3d, \simpeg) and linear interpolation (\custem, \petgem), in both cases on $\log_{10}{\rho}$ values. The actual representation of the original resistivities in the tetrahedra mesh is not as smooth as in the other meshes since this would require significantly smaller tetrahedra and more dof on top of the inclusion of subsurface layer constraints. However, note that the combination of slicing through tetrahedra and the vertical exaggeration lead to a very distorted view of the tetrahedral mesh which is barely representative for the overall averaged approximation of resistivities. The actual responses at the receivers (field interpolation) are obtained by spline interpolation (\emg3d), linear interpolation (\simpeg), and through the second-order basis functions (\custem, \petgem).

@@

%% 2. Discretization %%
One of the key insights we gained from this study is that creating appropriate meshes is a difficult and time-consuming task. While there exist now several open-source 3D CSEM codes the capability for automatic meshing lacks behind and currently remains a largely manual task requiring experience in the field. It is important to state that the chosen models favour structured meshes. The first two examples can easily and accurately be represented by regular meshes. Using, for instance, a dipping layer instead of three resistive cubes would turn it around and make the model more favourable for unstructured meshes. MR3D would be an ideal scenario for unstructured meshes, as it is based on also published irregular lithological horizons. Since the corresponding resistivity model, however, was defined on a regular grid, we were forced to approximate the comparatively fine regular discretization by an unstructured one and interpolate the resistivity data for being able to apply the FE codes to this problem. We are completely aware about the ineptness of this re-approximation procedure, unless it served for the cross-validation purposes. Both cases show that automatic and adaptive meshing capabilities are needed, which will increase the flexibility of 3D CSEM modelling. Such meshing codes should also take into account the physics of CSEM computation with its diffusive behaviour. It is also important to state that the unstructured mesh used for the MR3D model incorporates the entire domain and could be used for all source and receiver positions, whereas the rectilinear and octree meshes were designed for the shown receiver and corresponding source lines.

@@

We would like to thank Bane Sullivan for plotting of tetrahedra meshes in \texttt{matplotlib} through \texttt{PyVista}, Seogi Kang for the input and octree mesh design for the \simpeg layer and block models, and Joseph Capriotti for the octree mesh implementation and volume averaging in \simpeg. We would also like to thank Paulo Menezes for the help and explanations with regards to the Marlim R3D model and corresponding CSEM data, and for making their actual computation model available under an open-source license.
